Theories of deliberated judgment

### 2022-08-10 simple.tex
Proof sketch seems okay with def Axiom strengthened

Given $\ar \in \gargsi$, say that $\ar$ $i$-defends $\pga$ iff $\forall \hist \in \allhist \suchthat \ar \in \histend \land \hist \nileadsto \pga: d(\ar, \hist) ≠ \emptyset$.
Note that if an argument fails to $i$-defend $\pga$, then the argumentative strategy of the theory has failed: some path challenge an argument of the theory, and the theory is not ready to counter-argue against that path.

Accordingly, an argumentative theory claims that all the arguments required to convince $i$ of choosing $\pga$ do $i$-defend $\pga$.
It is said to be truthful iff this claim is correct.
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff $\forall i \in I, \ar \in \gargsi: \ar$ $i$-defends $\pga$.
\end{definition}

NB we can check $a_1, a_x, a_y, a_2, …$ but not other paths after $a_2$.
Thus, among the paths that contain $a_2$, they all contain $a_1$ before $a_2$.

@@

\begin{axiom}[Omitting the defended argument]
	\label{ax:omitting}
	If the theory is truthful, then $\forall \ar_1, \ar_2 \in \gargsi, \hist \suchthat \ar_2 \in \dgi(\ar_1) \land \ar_2 \in \histend: (\ar_1, \hist) \ileadsto \pga ⇒ \hist \ileadsto \pga$.
\end{axiom}
This is called an axiom because it cannot be falsified using observables. It has to be accepted on the basis of the rest of the observations. What it states are counterfactuals: if $i$ had been presented with $\hist$ not containing $\ar_1$, and assuming $i$ would have accepted an argument $\ar_2$ for $\pga$, …
This seems reasonable because $a_2$, as a defender of $a_1$, “contains” $a_1$. 
(To be finished.)

An argumentative theory is said to be correct iff $\forall i \in I: \pga \in \iprops$.
As previously, it is said to be valid iff its truthfulness leads to its correctness, except that the above axiom needs now be included as a price for the added flexibility of the theory.
\begin{theorem}[Validity]
	If an argumentative theory is truthful, then \cref{ax:omitting} implies that the theory is correct.
\end{theorem}
\begin{proof}
	Assume, to start with, that $\arga$ is not decisive because $\hist \nileadsto \pga, \arga \in \histend$.
	Assume that $\dgi(\arga) = \ar_2$ with no further hist against it.
	…
\end{proof}

>>>>

An argumentative theory claims that whenever some of its arguments fail to convince $i$ of choosing $\pga$, it can argue something else.
It is said to be truthful iff this claim is correct.
\begin{definition}[Truthfulness]
	\label{def:truth}
	An argumentative theory $(\pga, \arga, \dg)$ for the exposure protocol $(I, \allargs, {\leadsto})$ is \emph{truthful} iff $\forall i \in I, \ar_2 \in \dgi(\arga), \ar_3 \in \dgi(\ar_2), …, \forall \hist \in \allhist \suchthat \hist = (\arga, …, \ar_2, …, \ar_3, …) \land \ar_3 \in \histend: [\hist \nileadsto \pga ⇒ d(\ar_3, \hist) ≠ \emptyset]$.
\end{definition}

@@

\begin{axiom}[Omitting defended arguments]
	\label{ax:omitting}
	If the theory is truthful, then $\forall \ar_2 \in \dgi(\arga), \ar_3 \in \dgi(\ar_2), …, \hist \suchthat \ar_7 \in \histend: (…, \ar_4, \ar_5, …, \hist) \ileadsto \pga ⇒ (…, \ar_5, …, \hist) \ileadsto \pga$.
\end{axiom}
This is called an axiom because it cannot be falsified using observables. It has to be accepted on the basis of the rest of the observations. What it states are counterfactuals: if $i$ had been presented with $\hist$ not containing $\ar_1$, and assuming $i$ would have accepted an argument $\ar_2$ for $\pga$, …
This seems reasonable because $a_2$, as a defender of $a_1$, “contains” $a_1$. 
(To be finished.)

An argumentative theory is said to be correct iff $\forall i \in I: \pga \in \iprops$.
As previously, it is said to be valid iff its truthfulness leads to its correctness, except that the above axiom needs now be included as a price for the added flexibility of the theory.
\begin{theorem}[Validity]
	If an argumentative theory is truthful, then \cref{ax:omitting} implies that the theory is correct.
\end{theorem}
\begin{proof}
	Assume, to start with, that the theory is of depth 2, more precisely, $\arga$ is not decisive because $\hist \nileadsto \pga, \arga \in \histend$ and $\dgi(\arga) = \ar_2$ with no further hist against it.

	Consider any $i \in I$.
	Assume $\arga$ is not decisive.

	Truthfulness brings this.
	$\forall \hist = (\arga, …, \ar_2, …) \suchthat \ar_2 \in \histend: \hist \ileadsto \pga$.

	The axiom brings this.
	$\forall \hist \suchthat \ar_2 \in \histend: (\arga, \hist) \ileadsto \pga ⇒ \hist \ileadsto \pga$.

	Consider any $\hist \suchthat \ar_2 \in \histend$. We want to show that $\hist \ileadsto \pga$.
	If it starts with $\arga$: done thanks to truthfulness.
	Otherwise: done thanks to truthfulness and the axiom.

Consider now a theory of depth 3. 
	Consider any $i \in I$.
	Assume $\arga$ not decisive, $\ar_2$ not decisive.

	Truthfulness brings this.
	$\forall \hist = (\arga, …, \ar_2, …, \ar_3, …) \suchthat \ar_3 \in \histend: \hist \ileadsto \pga$.

	The axiom brings this.
	$\forall \hist \suchthat \ar_3 \in \histend: (\ar_2, \hist) \ileadsto \pga ⇒ (\hist) \ileadsto \pga$.
	And this.
	$\forall \hist \suchthat \ar_3 \in \histend: (\ar_1, \ar_2, \hist) \ileadsto \pga ⇒ (\ar_2, \hist) \ileadsto \pga$.
	And $(\ar_1, \ar_2, \hist) \ileadsto \pga$ holds thanks to truthfulness, so $(\ar_2, \hist) \ileadsto \pga$, so $\hist \ileadsto \pga$.

	Consider any $\hist \suchthat \ar_3 \in \histend$. We want to show that $\hist \ileadsto \pga$.
	If it includes a1 then a2, ok thanks to truthfulness.

\end{proof}
